MY RMD FILE FOR PA1
========================================================

## FIRST CLONE REPO FROM PENG
## this should create figures directory where figures should be placed in (automatically?)
## need to place Rmd in directory and then knit Rmd to get other files

# A Loading and preprocessing the data

## 1 Load and inspect the data


```{r, echo=F}
ip <- as.data.frame(installed.packages()[,c(1,3:4)])
rownames(ip) <- NULL
ip <- ip[is.na(ip$Priority),1:2,drop=FALSE]
print(ip, row.names=FALSE)
```

```{r}
getwd()
dir()
ls()

act<-read.csv("activity.csv")
head(act)
tail(act)
class(act)
str(act)
```

## 2 Process/transform the data

Turn dates into date class.
```{r}
act$date<-as.Date(act$date)
str(act)
```

Reorder the columns.
```{r}
act<-act[, c(2,3,1)]
head(act)
```

Create wide dataset for the analysis per 5-min interval (gcook p367)  ## check if needed
```{r}
install.packages("reshape2")
library("reshape2", lib.loc="C:/Users/erobesyn/Documents/R/win-library/3.0")
```


```{r, results="hide"}
actwide<-dcast(act, act$date~act$interval, value.var="steps")
str(actwide)
#View(actwide)
```
Actwide is dataframe of 61 rows (days) and 288 other columns (5-min intervals in one day)

[Convert dataframe in 2-col df with one for date with fractional values (gcook p 370)
(cave I do not have time series object yet)
zoo package to convert to time series object ?? ]

# B What is mean total number of steps taken per day?

## 1 Make histogram of total number of steps per day

1. First, we create a new variable, called "stepsbyday", to store the total number of steps per day. 
```{r}
stepsbyday<-tapply(act$steps, act$date, sum)
head(cbind(stepsbyday))
plot(stepsbyday, type="l")
```

2. Then, we make the histogram of this new variable.  
```{r,fig.width=7,fig.height=4}
hist(stepsbyday, col="red", main="Histogram of steps per day", ylab="Frequency", xlab="Number of steps per day", breaks=20, xlim=c(0,max+3000))
``` 

```{r date with max steps - not tasked}
max<-max(stepsbyday[is.na(stepsbyday)=="FALSE"])
max
which.max(stepsbyday[is.na(stepsbyday)=="FALSE"])   # in positie 47 
#View(stepsbyday[is.na(stepsbyday)=="FALSE"])       # confirm position 47 has max value 

# if I do not delete NA's first, I will find same max value but in position 54
# which.max(stepsbyday)
# View(stepsbyday)
```


## 2 Calculate mean and median total number of steps per day

```{r}
meanbyday<-mean(stepsbyday,na.rm=T)
medianbyday<-median(stepsbyday, na.rm=T)
meanbyday
medianbyday
```

The mean, accross all days, of the total number per day is `r meanbyday` and the median total number per day is `r medianbyday`.
They are only 1 step per day different from each other, indicating hardly any skew in the data.

Note: For some reason the correct mean and median calculated by the 5-number summary differs very slightly from the command summary(stepsbyday). Summary() finds the maximum value at 21190 while the real maximum value is 21194.  (rounding effect?)

The mean, and median, per day, is a dataframe with 61 values each

## above I worked with wide dataset actwide but dplyr works with long data (the original act dataframe)
```{r}
library(dplyr)
act<-tbl_df(act)
byday<-act %>% group_by(date) %>% summarize(mean= mean(steps,na.rm=T))
byday
plot(byday)
plot(byday, type="l", col="red")
```

[Store MEAN total steps per day and make ts of this var]  ??? NOT NEEDED ???


# C What is the average daily activity pattern?

# plot(act$date, act$steps, type="l")

## 1 Make time series plot with average number of steps per 5-min interval, averaged across all days
Do the same as above but with MEAN per 5-min interval

```{r}
library(dplyr)
act<-tbl_df(act)
byinterval<-act %>% group_by(interval) %>% summarize(mean= mean(steps,na.rm=T))
byinterval
plot(byinterval)
plot(byinterval, type="l", col="red")
```




```{r}
stepsperinterval<-colSums(actwide[,-1],na.rm=T)  # all columns except first col that contains dates
head(stepsperinterval)
tail(stepsperinterval)

``` 

First sum over all 61 days, then divide by 61 to get average.
```{r, fig.width=7, fig.height=4}
max(stepsperinterval)
max(stepsperinterval)/61 

png(file="testplot.png")
plot(stepsperinterval/61, type="l", main="Average number (across all days) of steps per 5-min interval", ylab="Number of steps", xlab="5-min intervals in a day", col="red")
dev.off()
```

## 2 Which 5-min interval, averaged across all days, contains most steps?


10927 steps in 5 min - to divide by 61 days? =179 steps per 5 min

```{r,results="hide"}
stepsperinterval<-as.data.frame(stepsperinterval)
stepsperinterval/61
```

```{r}
str(stepsperinterval/61)
```
There are 288 5-min intervals in a day

Look for interval with 179.13 steps
Answer: interval 835! (cumulative nr of 5-min intervals) 
WRONG because we are looking at averages across days...


# D Imputing missing values

## 1 Calculate total number of rows with NAs

```{r NA}
rows<-nrow(act)
narows<-sum(is.na(act))
```
From the `r rows` in the dataset, `r narows` contain one or more missing values.

## 2 Devise strategy for filling missing values (e.g. mean for day, mean for 5-min interval, ...)
## 3 Create new dataset with missing data filled in.
## 4 Make histogram with total nr of steps per day. Calculate mean and median total steps per day. Do they differ from part B (without imputing missing values). What is impact of imputing missing data on the estimates?


# E Are there differences in activity patterns between weekdays and weekends?

## 1 Create factor 'part of week' (weekday or weekend)
## 2 Make panel plot of time series with average number of steps per 5-min interval, averaged across all weekdays OR weekenddays.

See gcook p354. ddply with transform

Then ggplot2 with gcook p370
Or zoo rcook 

```{r}
library(zoo)
datets<-zoo(act,act$date)
str(datets)
head(datets)
```

