MY SCRIPT FOR PA1 REPORT
=======================================================

# A Loading and preprocessing the data

## 1 Load and inspect the data

Load the data from the csv file and look at the first and last rows and the data structure.
```{r}
setwd("C:/Users/Robesyn/RepData_PeerAssessment1/activity")
dir()
act<-read.csv("activity.csv")
head(act)
tail(act)
class(act)
str(act)
```

## 2 Process/transform the data

Turn dates into date class.
```{r}
act$date<-as.Date(act$date)
```

Reorder the columns, to have the date first, then the interval within the date, and then the number of steps for the corresponding date-time interval.
```{r}
act<-act[, c(2,3,1)]
head(act)
```

Create a wide dataset for the analysis per 5-min interval.
```{r}
install.packages("reshape2")
library("reshape2", lib.loc="C:/Users/Robesyn/Documents/R/win-library/3.0")
```

```{r, results="hide"}
actwide<-dcast(act, act$date~act$interval, value.var="steps")
str(actwide)
```
"Actwide" is a dataframe of 61 rows (days of activity measurement) and 288 other columns (the consecutive 5-min intervals in one day).

# B What is mean total number of steps taken per day?

## 1 Make histogram of total number of steps per day

1. First, we create a new variable, called "stepsbyday", to store the total number of steps per day. 
```{r}
stepsbyday<-tapply(act$steps, act$date, sum)
head(cbind(stepsbyday))
```

2. Then, we make the histogram of this new variable.  
```{r,fig.width=7,fig.height=4}
max<-max(stepsbyday[is.na(stepsbyday)=="FALSE"])
max

hist(stepsbyday, col="red", main="Histogram of steps per day", ylab="Frequency", xlab="Number of steps per day", breaks=20, xlim=c(0,max+3000))
``` 

## 2 Calculate mean and median total number of steps per day

```{r}
meanbyday<-mean(stepsbyday,na.rm=T)
meanbyday<-as.integer(meanbyday)
medianbyday<-median(stepsbyday, na.rm=T)
```

The mean total number per day is `r meanbyday` and the median total number per day is `r medianbyday`. They are only 1 step per day different from each other, indicating hardly any skew in the data.

# C What is the average daily activity pattern?

First plot the raw data per day.  Each dot is the number of steps in a 5 minutes interval on a specific day. 

```{r raw data}
plot(act$date, act$steps, pch=19, cex=0.0001, col="blue")
```

## 1 Make time series plot with average number of steps per 5-min interval, averaged across all days

```{r}
meanperinterval<-sapply(actwide[,-1],mean, na.rm=T)
plot.ts(meanperinterval,type = "l", main = "Average number (across all days) of steps per 5-min interval", ylab = "Number of steps", xlab = "5-min intervals in a day", col = "red")
```


## 2 Which 5-min interval, averaged across all days, contains most steps?


10927 steps in 5 min - to divide by 61 days? =179 steps per 5 min

```{r,results="hide"}
stepsperinterval<-as.data.frame(stepsperinterval)
stepsperinterval/61
```

```{r}
str(stepsperinterval/61)
```
There are 288 5-min intervals in a day

Look for interval with 179.13 steps
Answer: interval 835! (cumulative nr of 5-min intervals) 

# D Imputing missing values

## 1 Calculate total number of rows with NAs

```{r NA}
rows<-nrow(act)
narows<-sum(is.na(act))
```
From the `r rows` in the dataset, `r narows` contain one or more missing values.

## 2 Devise strategy for filling missing values (e.g. mean for day, mean for 5-min interval, ...)
Replace the NA values by mean per day, from section B.

## 3 Create new dataset with missing data filled in.

## 4 Make histogram with total nr of steps per day. Calculate mean and median total steps per day. Do they differ from part B (without imputing missing values). What is impact of imputing missing data on the estimates?


# E Are there differences in activity patterns between weekdays and weekends?

## 1 Create factor 'part of week' (weekday or weekend)

## 2 Make panel plot of time series with average number of steps per 5-min interval, averaged across all weekdays OR weekenddays.


```{r}
install.packages("zoo")
library("zoo", lib.loc="C:/Users/Robesyn/Documents/R/win-library/3.0")

datets<-zoo(act,act$date)
str(datets)
head(datets)
```
