---
title: "RepData_PeerAssessment1_v2"
author: "Robesyn"
date: "Thursday, June 11, 2015"
output: html_document
---

REPRODUCIBLE RESEARCH, PROJECT 1 - DAILY ACTIVITY MEASUREMENTS
===============================================================

## FIRST CLONE REPO FROM PENG
## this should create figures directory where figures should be placed in (automatically?)
## need to place Rmd in directory and then knit Rmd to get other files

# A Loading and preprocessing the data

## 1 Prepare, read and inspect
Packages
```{r packages installed, echo=F}
ip <- as.data.frame(installed.packages()[,c(1,3:4)])
rownames(ip) <- NULL
ip <- ip[is.na(ip$Priority),1:2,drop=FALSE]
print(ip, row.names=FALSE)
```

```{r install new packages if needed, echo=F}
```

```{r load packages, echo=F}
library(dplyr)
```

Workspace
```{r prepare workspace}
getwd()
dir()
ls()
```

Data
```{r read data}
act<-read.csv("activity.csv")
head(act)
tail(act)
class(act)
str(act)
```

## 2 Process/transform the data
Turn dates into date class.
```{r}
act$date<-as.Date(act$date)
str(act)
```

Reorder the columns.
```{r}
act<-act[, c(2,3,1)]
head(act)

act<-tbl_df(act)
```

# B What is mean total number of steps taken per day?

## 1. Create new variable, called "totalstepsbyday", to store the total number of steps per day. 
```{r total steps by day}
totalstepsbyday<-act %>% group_by(date) %>% summarize(sum= sum(steps))
totalstepsbyday

sum<-totalstepsbyday$sum
sum

plot(totalstepsbyday)
plot(totalstepsbyday, type="l", col="red", main="Total steps per day")
```

## 2. Histogram of total steps per day.  
```{r histogram total steps by day,fig.width=7,fig.height=4}
hist(sum, col="red", main="Histogram of steps per day", ylab="Frequency", xlab="Number of steps per day", breaks=20, xlim=c(0,max+3000))

# store figure
png(file="histogram of total steps by day.png")
hist(sum, col="red", main="Histogram of steps per day", ylab="Frequency", xlab="Number of steps per day", breaks=20, xlim=c(0,max+3000))
dev.off()
``` 

## 3. mean and median of total steps by day
```{r mean and median of total steps by day}
meanoftotalstepsbyday<-mean(totalstepsbyday$sum, na.rm=T)
medianoftotalsteps<-median(totalstepsbyday$sum, na.rm=T)
meanoftotalstepsbyday
medianoftotalsteps
```
The mean, accross all days, of the total number per day is `r meanoftotalstepsbyday` and the median total number per day is `r medianoftotalstepsbyday`.  They are only 1 step per day different from each other, indicating hardly any skew in the data.

```{r mean steps by day - not tasked}
meanstepsbyday<-act %>% group_by(date) %>% summarize(mean= mean(steps,na.rm=T))
meanstepsbyday
plot(meanstepsbyday)
plot(meanstepsbyday, type="l", col="red", main="Average number of steps per day")
```

# C What is the average daily activity pattern?

## 1 Make time series plot with average number of steps per 5-min interval, averaged across all days

Do the same as above but with MEAN per 5-min interval
```{r}
stepsbyinterval<-act %>% group_by(interval) %>% summarize(mean= mean(steps,na.rm=T))
stepsbyinterval
plot(stepsbyinterval)
plot(stepsbyinterval, type="l", lwd=2, col="red", main="All day mean number of steps by 5-minute interval")

png(file="mean number of steps by interval")
plot(stepsbyinterval, type="l", lwd=2, col="red", main="All day mean number of steps by 5-minute interval")
dev.off()
```

## 2. Which 5-minute interval contains maximum number of steps?
```{r interval with maximum number of steps}
# max value
stepsbyinterval
max<-max(stepsbyinterval$mean)
max

# position of max value
which.max(stepsbyinterval$mean)   # 104th 5-minute interval
stepsbyinterval$mean              # check: max value indeed in position 104
```

# D Imputing missing values

## 1 Calculate total number of rows with NAs

```{r NA}
rows<-nrow(act)
narows<-sum(is.na(act))
rows
narows
```
From the `r rows` rows in the dataset, `r narows` contain a missing value.

```{r alternative}
incompleterows<-act[!complete.cases(act),]    
dim(incompleterows)       
head(incompleterows)
tail(incompleterows)
```

## 2 Devise strategy for filling missing values (e.g. mean for day, mean for 5-min interval, ...)

### we chose to replace NA with mean for corresponding 5-minute interval  
act<-mutate(act, imputed=steps)
actimputed<- filter(act, steps==NA) %>% mutate(imputed=stepsbyinterval$mean)


actimputed<-filter(act, steps==NA) %>% group_by(interval) %>% mutate(imputed=stepsbyinterval$mean)
#%>% ifelse(steps=="NA", stepsbyinterval$mean, act$steps)


actimputed<-act %>% group_by(interval) %>% filter(act, steps==NA) %>% mutate(imputed=mean(act$steps, rm.na=T))

## 3 Create new dataset with missing data filled in.

## 4 Make histogram with total nr of steps per day. Calculate mean and median total steps per day. Do they differ from part B (without imputing missing values). What is impact of imputing missing data on the estimates?


# E Are there differences in activity patterns between weekdays and weekends?

## 1 Create factor 'part of week' (weekday or weekend)
## 2 Make panel plot of time series with average number of steps per 5-min interval, averaged across all weekdays OR weekenddays.
