---
title: "RepData_PeerAssessment1_v2"
author: "Robesyn"
date: "Thursday, June 11, 2015"
output: html_document
---

REPRODUCIBLE RESEARCH, PROJECT 1 - DAILY ACTIVITY MEASUREMENTS
===============================================================

## FIRST CLONE REPO FROM PENG
## this should create figures directory where figures should be placed in (automatically?)
## need to place Rmd in directory and then knit Rmd to get other files

# A Loading and preprocessing the data

## 1 Prepare, read and inspect
Packages
```{r packages installed, echo=F}
ip <- as.data.frame(installed.packages()[,c(1,3:4)])
rownames(ip) <- NULL
ip <- ip[is.na(ip$Priority),1:2,drop=FALSE]
print(ip, row.names=FALSE)
```

```{r install new packages if needed, echo=F}
```

```{r load packages, echo=F}
library(dplyr)
```

Workspace
```{r prepare workspace}
getwd()
dir()
ls()
```

Data
```{r read data}
act<-read.csv("activity.csv")
head(act)
tail(act)
class(act)
str(act)
```

## 2 Process/transform the data
Turn dates into date class.
```{r}
act$date<-as.Date(act$date)
str(act)
```

Reorder the columns.
```{r}
act<-act[, c(2,3,1)]
head(act)

act<-tbl_df(act)
```

# B What is mean total number of steps taken per day?

## 1. Create new variable, called "totalstepsbyday", to store the total number of steps per day. 
```{r total steps by day}
totalstepsbyday<-act %>% group_by(date) %>% summarize(sum= sum(steps))
totalstepsbyday

sum<-totalstepsbyday$sum
sum

plot(totalstepsbyday)
plot(totalstepsbyday, type="l", col="red", main="Total steps per day")
```

## 2. Histogram of total steps per day.  
```{r histogram total steps by day,fig.width=7,fig.height=4}
hist(sum, col="red", main="Histogram of steps per day", ylab="Frequency", xlab="Number of steps per day", breaks=20, xlim=c(0,max+3000))

# store figure
png(file="histogram of total steps by day.png")
hist(sum, col="red", main="Histogram of steps per day", ylab="Frequency", xlab="Number of steps per day", breaks=20, xlim=c(0,max+3000))
dev.off()
``` 

## 3. mean and median of total steps by day
```{r mean and median of total steps by day}
meanoftotalstepsbyday<-mean(totalstepsbyday$sum, na.rm=T)
medianoftotalsteps<-median(totalstepsbyday$sum, na.rm=T)
meanoftotalstepsbyday
medianoftotalsteps
```
The mean, accross all days, of the total number per day is `r meanoftotalstepsbyday` and the median total number per day is `r medianoftotalstepsbyday`.  They are only 1 step per day different from each other, indicating hardly any skew in the data.

```{r mean steps by day - not tasked}
meanstepsbyday<-act %>% group_by(date) %>% summarize(mean= mean(steps,na.rm=T))
meanstepsbyday
plot(meanstepsbyday)
plot(meanstepsbyday, type="l", col="red", main="Average number of steps per day")
```

# C What is the average daily activity pattern?

## 1 Make time series plot with average number of steps per 5-min interval, averaged across all days

Do the same as above but with MEAN per 5-min interval
```{r}
stepsbyinterval<-act %>% group_by(interval) %>% summarize(mean= mean(steps,na.rm=T))
stepsbyinterval
plot(stepsbyinterval)
plot(stepsbyinterval, type="l", lwd=2, col="red", main="All day mean number of steps by 5-minute interval")

png(file="mean number of steps by interval")
plot(stepsbyinterval, type="l", lwd=2, col="red", main="All day mean number of steps by 5-minute interval")
dev.off()
```

## 2. Which 5-minute interval contains maximum number of steps?
```{r interval with maximum number of steps}
# max value
stepsbyinterval
max<-max(stepsbyinterval$mean)
max

# position of max value
which.max(stepsbyinterval$mean)   # 104th 5-minute interval
stepsbyinterval$mean              # check: max value indeed in position 104
```

# D Imputing missing values

## 1 Calculate total number of rows with NAs

```{r NA}
rows<-nrow(act)
narows<-sum(is.na(act))
rows
narows
```
From the `r rows` rows in the dataset, `r narows` contain a missing value.

```{r alternative}
incompleterows<-act[!complete.cases(act),]    
dim(incompleterows)       
head(incompleterows)
tail(incompleterows)
```

## 2 Devise strategy for filling missing values (e.g. mean for day, mean for 5-min interval, ...)

We choose to replace NA with mean for corresponding 5-minute interval  

## 3 Create new dataset with missing data filled in.
actimputed<-transform(act, imputedsteps=ifelse(is.na(steps), ave(steps,interval, FUN=function(x) mean(x, na.rm=T)),steps))
actimputed<-tbl_df(actimputed)
View(actimputed)

# actimputed<-act %>% group_by(interval) %>% mutate(imputed=replace(steps, is.na(steps), mean(steps, rm.na=T)))  # why does this not work ???


## 4 Make histogram with total nr of steps per day. Calculate mean and median total steps per day. Do they differ from part B (without imputing missing values). What is impact of imputing missing data on the estimates?

totalimputedstepsbyday<-actimputed %>% group_by(date) %>% summarize(sum=sum(imputedsteps))
totalimputedstepsbyday

sumimputedsteps<-totalimputedstepsbyday$sum
sumimputedsteps

hist(sumimputedsteps, col="red", main="Histogram of steps, with imputation, per day", ylab="Frequency", xlab="Number of steps per day", breaks=20)

# save figure
png(file="histogram of total imputed steps by day.png")
hist(sumimputedsteps, col="red", main="Histogram of steps, with imputation, per day", ylab="Frequency", xlab="Number of steps per day", breaks=20)
dev.off()

# E Are there differences in activity patterns between weekdays and weekends?

Yes, they are different. This illustrates impact of imputation (and the method of imputation), which is important to be aware of.

## 1 Create factor 'part of week' (weekday or weekend)
## 2 Make panel plot of time series with average number of steps per 5-min interval, averaged across all weekdays OR weekenddays.
